---
title: "Math441-Project 1 - Note"
author: "Dr. Marie-Claire Koissi"
date: "Fall 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
subtitle: Example of Confidence Interval
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK 4: Example of Confidence / Prediction Interval

This is an example of calculation of confidence interval and prediction interval. We will first use the formula from the note, then we will use Rstudio function **predict**.
For this example, we will work with the boys data set and use ht2 as independent variable (Not necessarly what you should use in the project).

```{r , echo=TRUE}

#TASK 2
###################
# Full regression analysis
# data =BOYSGROWTHDATA.CSV 
#Descriptive stat
# x = ht2
# y =ht18

library(ggplot2)

BOYSGROWTHDATA<-read.csv("BOYSGROWTHDATA.CSV", header = TRUE)
attach(BOYSGROWTHDATA)
par(mfrow=c(1,1))
ggplot(BOYSGROWTHDATA, aes(x=ht2,y=ht18))+
  geom_point()

#Regression
mean(BOYSGROWTHDATA$ht2) 
lm1<-lm(ht18~ht2, data = BOYSGROWTHDATA)
summary(lm1)
```

## TASK 4 (1): Example of Confidence Interval

## TASK 4 (1a): Using Formaula to compute Confidence Interval
The $100(1-\alpha)\%$ Confidence Interval at the value $x=x^{*}$ is the following

$$ (\hat\beta_0 + \hat\beta_1 * x^{*} ) \pm t_(\alpha/2 ; df) * s\sqrt{\frac{1}{n}+\frac{(x^{*}-\bar{x})}{SS_{xx}}} $$
For the $95\%$ Confidence Interval for the mean height for 18-year-old boys with height 92 cm at age 2, we have $x^{*}=ht2 = 92$. The R-output gives the following values:
$$
\begin{aligned}
    \hat\beta_0 &= 79.9570 \\
    \hat\beta_1 &= 1.1205  \\
    n & = 66 \\  
    \\
    s &= \sqrt{\frac{SSE}{(n-2)}}=\sqrt{\frac{1861.30}{(66-2)}}=\sqrt{29.08}=5.3928  \\
    \\
    x^{*}-\bar{x} & = 92 - 88.37121
\end{aligned}
$$
The t-score $t_{(0.025 ; 64)}$ can be obtained using a t-table, a calculator [ invT(0.025, 64)], or R-studio as follows:

```{r}
(qt(.025,3))
```

```{r , echo=TRUE}
(SSxx<-sum(ht2^2)-66*(mean(ht2))^2)
```

$$
\begin{aligned}
    t_{(0.025 ; n-2)} & = t_{(0.025 ; 64)}  = -3.182446 \\
    SS_{xx} & = 716.5953
\end{aligned}
$$

Finally, we get the following $95 \% $ confidence interval

$$ 
\begin{aligned}
(\hat\beta_0 + \hat\beta_1 * x^{*}) & \pm t_(\alpha/2 ; df) * s\sqrt{\frac{1}{n}+\frac{(x^{*}-\bar{x})}{SS_{xx}}} \\
\\
(79.9570 + 1.1205 (92)) & \pm 3.1824 * 5.3928 \sqrt{\frac{1}{66}+\frac{(92 - 88.3712)}{716.5953}}
\\
183.043 & \pm 2.24
\end{aligned}
$$
```{r}
(lowernumber<-183.043 - 2.24)
(uppernumber<-183.043 + 2.24)
```
We get the following $95 \% $ confidence interval for the mean height
$$ (180.803, 185.283)$$
## TASK 4 (1b): Using R-studio function predict to compute Confidence Interval

```{r, echo=TRUE}
BOYSGROWTHDATA<-read.csv("BOYSGROWTHDATA.CSV", header = TRUE)
attach(BOYSGROWTHDATA)
lmt2<-lm(ht18~ht2, data = BOYSGROWTHDATA)
newdata<-data.frame(ht2=92)
predict.lm(lmt2, newdata, interval = "confidence", level = 0.95)
```
## TASK 4 (2): Computing Prediction Interval

## TASK 4 (2a): Using Formaula to compute Prediction Interval
The $100(1-\alpha)\%$ Prediction Interval at the value $x=x^{*}$ is the following

$$ (\hat\beta_0 + \hat\beta_1 * x^{*}) \pm t_(\alpha/2 ; df) * s\sqrt{1+\frac{1}{n}+\frac{(x^{*}-\bar{x})}{SS_{xx}}} $$
##TASK 4 (2b): Using R-studio function predict to compute Confidence Interval
```{r, echo=TRUE}
BOYSGROWTHDATA<-read.csv("BOYSGROWTHDATA.CSV", header = TRUE)
attach(BOYSGROWTHDATA)
lmt2<-lm(ht18~ht2, data = BOYSGROWTHDATA)
newdata<-data.frame(ht2=92)
predict.lm(lmt2, newdata, interval = "prediction", level = 0.95)
```