---
title: '1.3 HW: earthquake'
group: '4: Ashley and Maggie'
author: "4: Ashley and Maggie"
date: "2/11/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## open earthquake
```{r, echo = TRUE}
earthquake <- read.table("/Users/maggiewesterland/Desktop/school/441/datasets/EARTHQUAKE.txt", header = TRUE)
attach(earthquake)
names(earthquake)

```

minutes - X, magnitude - Y

## summary
```{r, echo = TRUE}
summary(MINUTE)

```
median > mean suggesting a left skewed distribution, however the numbers are close together so a normal distribution is implied

```{r, echo = TRUE}
summary(MAGNITUDE)

```
median > mean suggesting a left skewed distribution, however the numbers are close together so a normal distribution is implied

## boxplot
```{r, echo = TRUE}
boxplot(MINUTE, horizontal = TRUE)

```
The boxplot does not indicate any outliers for the minute variable.

```{r, echo = TRUE}
boxplot(MAGNITUDE, horizontal = TRUE)

```
The boxplot suggests a lot of variation in the magnitude with one possible outlier at 6.70.

## scatterplot
```{r, echo = TRUE}
plot(MAGNITUDE ~ MINUTE)

```
This is a scatterplot where minutes are denoted by the x-axis and magnitudes are denoted by the y-axis.

